Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-01 inst/paper/main.tex
stuff

@@

In this section we turn our attention to the expression passed for
evaluation to \eval. The \eval function accepts all types of R objects
as inputs, but in the majority of cases
(\AllExpressionInputEvalCallPerc), it is called with expressions. The
remaining \AllValueInputEvalCallPerc are values, \ie, they evaluate to
themselves.
%
Passing a value into an eval is effectively an no-op so we look
closely at the expression types the call sites pass to eval across all
the runs. \AllValueInputEvalSitePerc of all call sites always pass
values, \AllExpressionInputEvalSitePerc pass expressions and a very
large proportion, \AllPolymorphicInputEvalSitePerc, pass both
expressions and values as inputs across all program runs. We inspected
the eval callsites that accept only values or both values and
expression as inputs. The top three of these call sites contribute to
over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this
function finds the default values for parameters with those names in
the caller function. Since the default value for a function parameter
can be a complex expression or a benign value such as \c{NULL}, the
\c{match.arg} function evaluates it through a call to \eval. This
makes the \eval inside \c{match.arg} ``polymorhpic''. The second such
\eval call originates from \ggplot, a popular R package for making
plots. \ggplot has it's own object system, \ggproto, which uses \eval
inside the \ggproto function to determine the super-class from which
it inherits (the default value is \c{NULL} but in general it could be
an arbitrarily expression). The third \eval call is in core R \c{str}
function used to print a summary of R objects. It internally maps
\eval function on the bindings of an environment which results in
inputs of multiples types being passed to \eval.

We further investigate the type of input passed to \eval in
Figure~\ref{fig:eval-expression-kind} by the type of input and the
source of \eval. In this figure, \emph{Expression} is the type of
program text, \emph{Environmment} is the type of scopes or
environments used for binding, \emph{Null} is the type of the unique
\c{NULL} object, \emph{Vector} is the type assigned to vectors of
strings, booleans, integers, reals and bytes; \emph{Symbol} is the
type of unevaluated symbols, and \emph{Function} is the type of
functions.

From the figure, we observe that majority of inputs to Core \eval
functions are expressions in contrast to package \eval calls which
receive mostly environments as inputs. In fact,
\PackageEnvironmentInputEvalCallPerc of \eval calls in packages
receive environments as input expressions. There are only four
callsites in package evals that receive environments as expressions.
\begin{compactitem}[$-$]
\item \c{ggplot2::ggproto} contributes to 99.9\% of package \eval
  calls that receive environments as input.
\item \c{R6::generator_funs} function contributes to only 936 \eval
  calls and implements the same functionality as \c{ggplot2::ggproto}
  but for the \emph{R6} OOP system.
\item \c{future::backtrace} function applies \eval to a future object
  which is implemented as an environment. This is called only once.
\item \c{RModel::str.RMmodel} function overloads the core \c{str}
  method for its \c{RModel} objects and maps \eval exactly like
  \c{str}. This is called only once.
\end{compactitem}

>>>>

\mypara{Eval Arguments.} Figure~\ref{fig:eval-expression-kind} breaks
down the expressions passed to \eval by R data type. It separates the
core packages from CRAN as they have different characteristics. In
core, XXX036\% of arguments in core are expressions. Whereas, in CRAN
package XXX037\% are expressions. When the argument is not an
expression, it is a value. Values are trivial for \eval as they are
returned unchanged. \PackageEnvironmentInputEvalCallPerc of CRAN
arguments are environments. This is an anomaly as only four \eval
call sites dominate this result:
\begin{compactitem}[$-$]
\item \c{ggplot2::ggproto} contributes to 99.9\% of package \eval
  calls that receive environments as input.
\item \c{R6::generator_funs} function contributes to only 936 \eval
  calls and implements the same functionality as \c{ggplot2::ggproto}
  but for the \emph{R6} OOP system.
\item \c{future::backtrace} function applies \eval to a future object
  which is implemented as an environment. This is called only once.
\item \c{RModel::str.RMmodel} function overloads the core \c{str}
  method for its \c{RModel} objects and maps \eval exactly like
  \c{str}. This is called only once.
\end{compactitem}

@@
